---
title: "10.2 Exercise"
author: "Justin Wisniewski"
date: '2022-05-21'
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Instructions:

**Fit a binary logistic regression model to the data set that predicts whether or not the patient survived for one year (the Risk1Y variable) after the surgery. Use the glm() function to perform the logistic regression. See Generalized Linear Models for an example. Include a summary using the summary() function in your results.**

```{r echo=TRUE, include=TRUE}
setwd("C:/Users/jwiz3/Desktop/Data Statistics/dsc520")
library('foreign')

thoracicSurgery_df <- read.arff("data/ThoraricSurgery.arff")

#Logistic Regression Model
thoracicSurgery_glm <- glm(Risk1Yr ~ DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 + PRE30 + PRE32 + AGE, data=thoracicSurgery_df, family = binomial)

summary(thoracicSurgery_glm)
```

**According to the summary, which variables had the greatest effect on the survival rate?**

```
    The following variables had the greatest effect on the survival rate (based on P value): 
    1. PRE9T - Indicates whether the patient had Dyspnoea before surgery.
    2. PRE14OC14 - The largest size of the original tumor.
    3. PRE17T - This variable indicates whether the patient had Type 2 Diabetes.
    4. PRE30T - Indicates that patient is a smoker. 
    5. PRE14OC13 - The second largest size of the tumor.
    6. PRE5 - Volume that has been exhaled at the end of the first second of forced expiration.
```

**To compute the accuracy of your model, use the dataset to predict the outcome variable. The percent of correct predictions is the accuracy of your model. What is the accuracy of your model?**

```
The accuracy of the model is 83.62%, so we can conclude that our model is correct in predicting the outcome.
```

```{r echo=TRUE, include=TRUE} 
res_val <- predict(thoracicSurgery_glm, type="response")
surgPredictionData <- table(Actual_Value = thoracicSurgery_df$Risk1Yr, Predicted_Value = res_val > 0.5)
surgPredictionData
modelAccuracy <- (surgPredictionData[[1,1]] + surgPredictionData[[2,2]]) / sum(surgPredictionData)
modelAccuracy
```
